---
title: "HeatMap"
author: "YiLin Liu"
date: "April 11, 2017"
output: 
    html_document:
    toc: true
    toc_depth: 6
---

##  Learning How To:

*   Main library: [rmaps](http://rmaps.github.io/).

*   `rmaps` is a R wrapper to interface between
    some popular Javascript packages and R. The packages
    mentioned are:
    +   [Crosslet](http://sztanko.github.io/crosslet/)
    +   [dataMaps](http://datamaps.github.io/)
    +   [leaflet](http://leafletjs.com/)

*   The heat map will be an overlay on top of a leaflet 
    interactive heat map.
    
*   An example is provided on [heatmap blog](http://rmaps.github.io/blog/posts/leaflet-heat-maps/index.html)
    +   uses the [leaflet.heat plugin](https://github.com/Leaflet/Leaflet.heat)


##  Following the [Example](http://rmaps.github.io/blog/posts/leaflet-heat-maps/index.html):

*   The data produced from the `Get Data` script is used with 
    this example. 
    
1: libraries

```{r, warning=F, message = F}

# require(devtools)
# install_github('ramnathv/rCharts@dev')
# install_github('ramnathv/rMaps')
```


2: Example:

*   Set up basic Map:


```{r load map}
    rm(list = ls())
    library(rMaps)
    library(rCharts)
  ## using the produced data for cities
    dataPath <- ".\\Deliverables\\dataComplete.csv"
    dataSet <- read.table(dataPath, header=T, sep=",")
    

    L2 <- Leaflet$new()
    L2$setView(c(median(dataSet$lat),  median(dataSet$lng)), 6)
    L2$tileLayer(provider = "OpenStreetMap.Mapnik")
    L2

```


*   Data:
    +   obtained form crime data form ggmap package.
    
```{r}
    # install.packages('ggmap')
    library(ggmap)
     library(rCharts)
     library(plyr)

    
    scaleFactor <-1
       
    ## the first step is to divide up the crime data (ddply) by
    ##  lat, long and the produce a different data frame via "summarise"
    ##  that contains the lat,lng and count (scaled version of freq).
     dataSet_dat = ddply(dataSet, .(lat, lng), summarise, count = freq*scaleFactor)
     head(dataSet_dat,n=3)
    
    ## produces JSON like nested list(array of objects), 
    ## json: specifies if data is already in json.
     
    dataSet_dat = rCharts::toJSONArray2(na.omit(dataSet_dat), json = F, names = F)
    head(dataSet_dat,n=2 )
    
    ##  showing the json styled dataSet.
    cat(rjson::toJSON(dataSet_dat[1:2]))
```



*   plotting the maps:
    *   Requires decent understanding of JavaScript. The author is
        not familar with javascript. 
    
    *   The author has not been able to locate the documentation
        for the methods used below.
        [classes R](http://adv-r.had.co.nz/OO-essentials.html)
        tutorial of `RC` object system is useful.
        
    *   Did not find a way to access source code or documentation
        to the methods used.
```{r}

# Add leaflet-heat plugin. Thanks to Vladimir Agafonkin
L2$addAssets(jshead = c(
  "http://leaflet.github.io/Leaflet.heat/dist/leaflet-heat.js"
))

# Add javascript to modify underlying chart
L2$setTemplate(afterScript = sprintf("
<script>
  var addressPoints = %s
  var heat = L.heatLayer(addressPoints).addTo(map)           
</script>
", rjson::toJSON(dataSet_dat)
))

L2

library(pryr)
otype(L2)

```



##  R Kernel Density Functions:
[R evaluation of Kernel Package ](http://vita.had.co.nz/papers/density-estimation.pdf)

[Kernel Introduction](https://people.cs.pitt.edu/~milos/courses/cs3750-Fall2007/lectures/class-kernels.pdf)