---
title: "Rough Density Plot"
author: "YiLin Liu"
date: "April 11, 2017"
output: 
    html_document:
    toc: true
    toc_depth: 6
---


##  Rough Visulization
    
1: libraries

```{r, warning=F, message = F}
# 
# require(devtools)
# install_github('ramnathv/rCharts@dev')
# install_github('ramnathv/rMaps') 
# install.packages('ggmap')
# install.packages('pryr')

    rm(list = ls())

    library(stam)
    library(rMaps)
    library(rCharts)
    library(ggmap)
    library(rCharts)
    library(plyr)
    library(pryr)
```



2: preparing data:
*   STAM package, the ygrid is the columns, xgrid is the rows 
*   xpoints are long, ypoints are lat

```{r preparing data}
 
    ## file paths
    pointsPath <- "points.csv"
    kernalPath <- "300SamplesWith6bins"
    
   
    
    ## loading data
    points <- read.csv(pointsPath)
    kernalData <- readRDS(kernalPath)
    testData <- kernalData$dens[,,1]

    ##  Making the correct points:
    xincr <- (points$xpoints[11]-points$xpoints[1])/9
    yincr <- (points$ypoints[11]-points$ypoints[1])/9
    
    xpoints <- seq(points$xpoints[1],points$xpoints[11],xincr)
    ypoints <- seq(points$ypoints[1],points$ypoints[11],yincr)
    
    points <- data.frame(xpoints,ypoints)
    
    ## preparing data frame of lat,lng and density (height)
        ## place holder lists
        lat <- c()
        long <- c()
        density <- c()
        index <- 1
        
    
        ## navigating the testData by rows
        ## start at 1st row, then go across, then go
        ## to second row...
        
        for( i in 1: dim(testData)[1]){
            for(j in 1:dim(testData)[2]){
                
                density[index] <- testData[i,j]
                lat[index] <- points$ypoints[j]
                long[index] <- points$xpoints[i]
                ##increasing index 
                index <- index+1
            }
        }
    
        formattedData <- data.frame(lat,long,density)
        
        plot(x=formattedData$lat,y = formattedData$long,"p")
        
        
        ## simple colour contour
        contour(x=points$xpoints,y=points$ypoints,
                       z=testData)
    
    
```





2: Example:

*   Set up basic Map:


```{r load map}
  
  ## using the produced data for cities
    dataSet <- formattedData
    

    L2 <- Leaflet$new()
    L2$setView(c(median(dataSet$lat),  median(dataSet$long)), 6)
    L2$tileLayer(provider = "OpenStreetMap.Mapnik")
    L2

```


*   Data:
    +   obtained form crime data form ggmap package.
    
```{r}
   

    dataSet_dat = ddply(dataSet, .(lat, long), summarise, count = density)
    
    ## produces JSON like nested list(array of objects), 
    ## json: specifies if data is already in json.
    dataSet_dat = rCharts::toJSONArray2(na.omit(dataSet), json = F, names = F)
    head(dataSet_dat,n=2 )
    
    ##  showing the json styled dataSet.
    cat(rjson::toJSON(dataSet_dat[1:2]))
```



*   plotting the maps:
    *   Requires decent understanding of JavaScript. The author is
        not familar with javascript. 
    
    *   The author has not been able to locate the documentation
        for the methods used below.
        [classes R](http://adv-r.had.co.nz/OO-essentials.html)
        tutorial of `RC` object system is useful.
        
    *   Did not find a way to access source code or documentation
        to the methods used.
```{r}

# Add leaflet-heat plugin. Thanks to Vladimir Agafonkin
L2$addAssets(jshead = c(
  "http://leaflet.github.io/Leaflet.heat/dist/leaflet-heat.js"
))

# Add javascript to modify underlying chart
L2$setTemplate(afterScript = sprintf("
<script>
  var addressPoints = %s
  var heat = L.heatLayer(addressPoints).addTo(map)           
</script>
", rjson::toJSON(dataSet_dat)
))

L2



```



##  R Kernel Density Functions:
[R evaluation of Kernel Package ](http://vita.had.co.nz/papers/density-estimation.pdf)

[Kernel Introduction](https://people.cs.pitt.edu/~milos/courses/cs3750-Fall2007/lectures/class-kernels.pdf)