---
title: "Twitter Stream API"
author: "paul"
date: "March 20, 2017"
output:
    html_document:
        toc:true
        toc.depth: 6

    pdf_document:
        toc:true
        toc.depth: 6
---

```{r Stream API Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*****

### Summary of streamR:
*   provides easy access to the twitter stream
    API.
    
*   Handles connection, parsing, disconnecting,
    reconnecting, backing off and 
    writing to a file all in the background

### IMPORTANT NOTE:
*   __Make sure that the twitter application has "obb" specified
in the call back URL, this will take the user to the authorization 
page to extract the pin to set up twitter handshake.__


*   __It is recommended to store the RAW JSON/XML response into a file
    and then process it later on to reduce possible delay for
    streams.__

*   __The streaming functions provided by streamR only stores
    complete tweets and disregards deletion, updates, incomplete 
    posts etc.__

*   __User stream returns only data for the authenticated user 
    for this session. Which is the twitter test account for this
    report. Not much information due to the nature of the account
    being a test account.__


### libraries
*   [streamR:](https://cran.r-project.org/web/packages/streamR/index.html)
    Handles connecting and extracting information from twitter stream apis.
```{r Stream API libraries, include=FALSE }
    rm(list = ls())
    library(streamR)
    library(ROAuth)
    library(twitteR)
    library(httr)

```



*****


### Stream Example 1: Public Streams:

Setting up Token: Commented out because it is stored as a file.

```{r Stream Example 1, include=FALSE, echo=FALSE}

    require(streamR)
    require(ROAuth)
    require(twitteR)
    require(httr)
    

   ## Taken from examples in streamR Package Documentation



    ##   Set up Token, only needs to be run once, The token will 
    ##   be saved to a file for future use. 

    ##  Code for setting up handshake, commented out because token is saved
    ##  into a file

    # requestURL <- "https://api.twitter.com/oauth/request_token"
    # accessURL <- "https://api.twitter.com/oauth/access_token"
    # authURL <- "https://api.twitter.com/oauth/authorize"
    # consumer_Key = "oQ3PqERg75kPtgBcgOLaFShSC"
    # consumer_Secret = "d4cxaKc1Dt3ugagruUNPtWzvmqGHx8WvwYAQ8MywUqTIVTTj9O"
    # 
    # my_oauth <- OAuthFactory$new(consumerKey=consumer_Key,
    # consumerSecret=consumer_Secret, 
    # requestURL=requestURL,accessURL=accessURL, authURL=authURL)
    # 
    # my_oauth$handshake(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl"))
```


Loading access token file:

```{r Stream Exapmle 1: Saving and Loading Token}

    ##  The token is stored and read as an R object

    if(!file.exists("stream Token")){
        file.create("stream Token")
        saveRDS(object = my_oauth, file="stream Token")
       
    }
    
    token = readRDS("stream Token")
    
```


Streaming form stream API:

*   The streamed data is saved under `tweets_CNN.json`.

```{r stream Example 1: streaming}
    
    ##Creating a file to store data
    if(!file.exists("tweets_CNN.json")){
        file.create("tweets_CNN.json")
        
         ## Can be controlled by either number of tweets
        ## and maximum connection time (timeout)
    
        filterStream( file.name="tweets_CNN.json",
        track="CNN", tweets=10, oauth=token)
    }
     
    ## reading in saved file, convert it to a data frame.
    ## where each column is a field and each row is a tweet.
    tweets_DB = parseTweets(tweets = "tweets_CNN.json")
    names(tweets_DB)
    
    
    ## a list where each element is a JSON nested 
    ## tweet
    tweets_List = readTweets(tweets="tweets_CNN.json")

```


*****



### Stream API Example 2: UserStream

*   The streamed data is saved under `user stream.json`.

```{r Stream Example 2}
    ##  Not much information is provided beacuse
    ##  not many activites of the test account. 

    if(!file.exists("user stream.json")){
        file.create("user stream.json")
        userStream(file.name="user stream.json",
               timeout=120,oauth=token)
      
    }


```