---
title: "Open Street Map"
author: "paul"
date: "March 28, 2017"
output: 
    html_document:
        toc: true
        toc.depth: 5
---

*****

##  Websites Visited:
*   [Open Street Maps Wiki](http://wiki.openstreetmap.org/wiki/Comparision_Google_services_-_OSM)
*   [Open Street Wiki](http://wiki.openstreetmap.org/wiki/Main_Page)
*   [Open Street Database Access](http://wiki.openstreetmap.org/wiki/Databases_and_data_access_APIs)


*****

##  Testing Accounts:

*   MapQuest:
    +   _Username:_ NRC-API-Testing
    +   _Email address:_ nrc.api.testing@gmail.com
    +   _First Name:_ NRC
    +   _Last Name:_ Testing
    +   _Company:_ NRC
    +   _Password:_ NRCTesting123
    

*****

##  Quick Comparison Between Google and OSM (Overpass):

*   Google Maps API offers multiple services oriented to replicate
    the features of google maps. The calculations such as routing, 
    height features are unique and is not available directly through
    OSM database.

*   Google is a centeralized collection of user features that 
    can be accessed by the user. On the otherhand, OSM is a open and 
    collabrative data base of geographical data which has enabled 
    3rd parties to develop amazing features not directly owned 
    by OSM but is based on OSM.

*   From initial readings on the OSM overpass api reveals that
    it supports a flexible and powerful query language which gives
    much more control to extract specific collection of points.
        +   EX: to get all roads 
        that start with _k_ in the general location of a bounding box 
        and all the points along this road that is a shop.

*   OSM community supports development and has support for many features
    made available via other 3rd party services that often exptends on 
    what is available to the public via Google Maps API. 
    +   [3D building modelling](http://wiki.openstreetmap.org/wiki/3D#Exporting)
    +   [Routing](http://wiki.openstreetmap.org/wiki/Routing)
        +   [Map Features](http://wiki.openstreetmap.org/wiki/Map_Features)
        +   [OSM Route Tags](http://wiki.openstreetmap.org/wiki/OSM_tags_for_routing)
        +   [Open Source Routing Machine](http://project-osrm.org/)
            +   Provides an API for routing information.
    
    
*****

##  API/Data base Acess:
*   Weekly update of all raw data for the entire planent
    at [Planet.osm](http://wiki.openstreetmap.org/wiki/Planet.osm)

*   Not as much information as google Maps, a lot of utilities
    are not included in the database.
    Instead I will spend more time looking into features based 
    on open street maps.

*   Large Regions through the [Xapi](http://wiki.openstreetmap.org/wiki/Xapi)
    +   deals with only current data.
    +   based on OSM API, but more optimized.
    +   3 types of data:
        +   Nodes: individual points
        +   Ways: a collection of nodes which illustrates
                  a road/way of some sort.
        +   relations: organized node by a common 
            theme
*   Offical [OSM API](http://wiki.openstreetmap.org/wiki/API)
    +   mainly used for editing the online database

*   [Overpass API](http://wiki.openstreetmap.org/wiki/Overpass_API)
    +   more optimized for extraction of data. READ ONLY
    +   Opens up a lot more customizabilty compared to 
        google maps.
    +   [Overpass Language Guide](http://wiki.openstreetmap.org/wiki/Overpass_API/Language_Guide)
        +   Please read for the query structure.
    +   [Overpass R Package](https://github.com/hrbrmstr/overpass)
    +   [Geocoding Services OSM Package](https://github.com/hrbrmstr/nominatim)
        +   translating from nodes to lat/lng and meta information and vice versa.

*   For interest [Database Schema](http://wiki.openstreetmap.org/wiki/Databases_and_data_access_APIs)
    +   [Data Table Schema](http://wiki.openstreetmap.org/wiki/Rails_port/Database_schema)

*   [Rendering](http://wiki.openstreetmap.org/wiki/Rendering)

*   [Topographical Data]()

*   [fire station information](http://wiki.openstreetmap.org/wiki/OpenFireMap)

*****

##  Data Extraction of OSM DataBase:
__GOAL: Explore the basics of what the query language 
    of the [Overpass API](http://wiki.openstreetmap.org/wiki/Overpass_API/Language_Guide). 
    Overpass API is focused on reading the existing data. For developers
    who like to modify the existing data, use the [OSM API](http://wiki.openstreetmap.org/wiki/API).__

*   [Overpass R Package](https://github.com/hrbrmstr/overpass)
    +   provides functionality to access the Overpass API
        directly through R.
    +   [Overpass Query Tester](http://overpass-turbo.eu/)

*   [OSM Nominatim R](https://github.com/hrbrmstr/nominatim)
    +   provides geocoding services for lat/lng to data structures
        by name in OSM and vice versa.
    +   [nominatim online](https://nominatim.openstreetmap.org/)
    +   [nominatim api](http://wiki.openstreetmap.org/wiki/Nominatim)
    

```{r Library setup}

    rm(list = ls())
    library(nominatim)
    library(overpass)
    library(ROAuth)

```

1:  Trying out Geocoding:


```{r geocodingOSM}
    require(nominatim)
    require(ROAuth)

#   Need an API Key, not documented in the
#   help for the methods

    # creating key
    consumer_Key = "27PLwRimfDaCgDPUsIWT2hG8TyxJ7CaO"
    consumer_Secret = "QeTEqoWepPiNjSV4"


    ## search by address
    ## can not find tim hortons at gloucester center, not in the maps.
    ## lack of information will cause problems.
    query = "Tim Hortons"
    address_Response = osm_search(query=query,
        address_details=TRUE,key=consumer_Key,limit=5)
    head(address_Response, n=3)

    
    ## Lat and Lng


```
