---
title: "Twitter Stream API"
author: "paul"
date: "March 20, 2017"
output:
    html_document:
        toc:true
        toc.depth: 6

    pdf_document:
        toc:true
        toc.depth: 6
---

```{r Stream API Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*****


### IMPORTANT NOTE:
*   __Make sure that the twitter application has "obb" specified
in the call back URL, this will take the user to the authorization 
page to extract the pin to set up twitter handshake.__


*   __It is recommanded to store the RAW JSON response into a file
    and then process it later on to reduce possible delay for
    streams.__

*   __ The streaming functions provided by streamR only stores
    complete tweets and disregardes deletion, updates, incomplete 
    posts etc.__

*   __User stream returns only data for the authenticated user 
    for this session. Which is the twitter test account for this
    report.Not much information due to the nature of the account
    being a test account.__


### libraries
*   [streamR:](https://cran.r-project.org/web/packages/streamR/index.html)
    Handles connecting and extracting information from twitter stream apis.
```{r Stream API libraries, include=FALSE }
    rm(list = ls())
    library(streamR)
    library(ROAuth)
    library(twitteR)
    library(httr)

```



*****


### Stream Example 1: Public Streams:

```{r Stream Example 1, include=FALSE, echo=FALSE}

    require(streamR)
    require(ROAuth)
    require(twitteR)
    require(httr)
    

   ## Taken from examples in streamR Package Documentation



    #   Set up Token, only needs to be run once, The token will 
    #   be saved to a file for future use. 

    ##  Code for setting up handshake, commented out because token is saved
    ## into a file

    # requestURL <- "https://api.twitter.com/oauth/request_token"
    # accessURL <- "https://api.twitter.com/oauth/access_token"
    # authURL <- "https://api.twitter.com/oauth/authorize"
    # consumer_Key = "oQ3PqERg75kPtgBcgOLaFShSC"
    # consumer_Secret = "d4cxaKc1Dt3ugagruUNPtWzvmqGHx8WvwYAQ8MywUqTIVTTj9O"
    # 
    # my_oauth <- OAuthFactory$new(consumerKey=consumer_Key,
    # consumerSecret=consumer_Secret, 
    # requestURL=requestURL,accessURL=accessURL, authURL=authURL)
    # 
    # my_oauth$handshake(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl"))
```


```{r Stream Exapmle 1: Saving and Loading Token}

    if(!file.exists("stream Token")){
        file.create("stream Token")
        saveRDS(object = my_oauth, file="stream Token")
       
    }
     
    token = readRDS("stream Token")
    
```



```{r stream Example 1: streaming}
    
    ##Creating a file to store data
    if(!file.exists("tweets_CNN.json")){
        file.create("tweets_CNN.json")
        
         ## Can be controlled by either number of tweets
        ## and maximum connection time (timeout)
    
        filterStream( file.name="tweets_CNN.json",
        track="CNN", tweets=10, oauth=token)
    }
     
    ## reading in saved file, converted to a data frame.
    ## where each column is a field and each row is a tweet.
    tweets_DB = parseTweets(tweets = "tweets_CNN.json")
    names(tweets_DB)
    
    
    ## a list where each element is a JSON nested 
    ## tweet
    tweets_List = readTweets(tweets="tweets_CNN.json")

```


*****



### Stream API Example 2: UserStream

```{r Stream Example 2}
    
    if(!file.exists("user stream.json")){
        file.create("user stream.json")
        userStream(file.name="user stream.json",
               timeout=120,oauth=token)
      
    }


```