---
title: "Google Maps Mini-Project"
author: "paul"
date: "March 22, 2017"
output: 
    html_document:
    toc: true
    toc.depth: 5
---

*****

###  Description:
*   Extract 100 origin/destination city pairs. Where
    the origin are random cities in Alberta and
    the destination are cities outside of Alberta
    but inside Canada.
    +   To ensure reproducibility, the randomly 
        selected origin/destination pairs are stored
        under the csv file "progress.csv".
        
*   Store the addresses for origin/destination pairs
    along with their lat/long as well as route distance,
    time, average speed and straight line distance into
    a csv file.

*   Plot the origin/destination pairs with their route 
    onto a map.


*****


### Setting up List of Cities

*   [List of cities](http://www.canada-city.ca/all-cities-in-canada.php) copied into a text
    document named `list of cities`,
    needs reformatting to extract __city name__ and __province__.

```{r Setting up List of Cities For Extraction}
    ## Sample city raw format: 1: 100 Mile House, British Columbia 

    raw_Citys = read.table("List of Cities.txt",header = TRUE, sep ="\n")

    ## getting rid of numbers:
    
    raw_Citys = lapply(raw_Citys, function(x){
        gsub(pattern = "[0-9]+:", "",x)
    })
    
    raw_Citys = unlist(raw_Citys)
    
    
    ## Separating into city and province
    citys = list()
    province = list()
    for(i in 1: length(raw_Citys)){
        separated = unlist( strsplit(raw_Citys[[i]],",") )
        citys[[i]] = gsub(" ","",separated[1])
        province[[i]] = gsub(" ","",separated[2])
        
    }
    
    raw_Citys_DF = data.frame(city=unlist(citys),
                    province=unlist(province))

    rm(list = setdiff(ls(),c("raw_Citys_DF")))
```


*****

### Getting the orgin and destination cities randomly:

*   Randomly select origin and destination pairs.

*   The origin/destination pairs are stored under 
    "progress.csv". 

```{r orgin and destination}
    library(dplyr)


    ab_Citys = dplyr::filter(raw_Citys_DF, province=="Alberta")
    
    ## Selecting origin and dest citys
    origin_Citys = dplyr::sample_n(ab_Citys,100)
    dest_Citys = dplyr::sample_n(dplyr::filter(raw_Citys_DF,province != "Alberta"), 
                                 size=100)
    
    ## renaming variables 
    colnames(origin_Citys) = c("orig_city", "orig_province")
    colnames(dest_Citys) = c("dest_city","dest_province")
    
    ## formatting to out_DF and clearing memory
    out_DF = cbind(origin_Citys,dest_Citys)
    
    rm(list = setdiff(ls(),c("out_DF")))
    
    
    ## write the current out_DF to a file 
    ## to remove the random elements 
    # 
    # if(!file.exists("progress.txt")){
    #     file.create("progress.txt")
    # }
    # 
     # write.table(out_DF, file="progress.txt",sep=",",row.names=FALSE)

```


*****

### Getting Direction information


```{r google credentials}
    api_Key = "AIzaSyBGTs-gZCbyP8n0Hvw_VZ76Z6YrST1DNa8"
    host = "https://maps.googleapis.com/maps/api"

```


*   A function that uses the google maps API
    to find lat/lng for the cities.

```{r function for lat and lng}

## takes lists containing: (orig_City, orig_Province,
##  dest_city,dest_provice) and a api_Key
##  assume that the lists are matching and same length

## return a data frame containing lat,lng, route information

    lat_Lng = function(orig_city,orig_province,
                       dest_city,dest_province,
                        api_Key){
        
        ##libraries needed
        require(jsonlite)
        require(httr)
        
        
        ## some google request information
        host = "https://maps.googleapis.com/maps/api"
        service = "/directions/json"
        
        
        ## lists to fill for a data frame
        orig_lat = list()
        orig_lng = list()
        orig_address = list()
        dest_lat = list()
        dest_lng = list()
        dest_address = list()
        polyline = list()
        time_text = list()
        time_value = list()
        distance_text = list()
        distance_value = list()
        
        
        ##looping through list and get the lat/lng
        for( i in 1:length(orig_city)){
            
            ## Getting names of city and province
            origin = paste(orig_city[[i]],orig_province[[i]],sep=",")
            origin = gsub(" ", "+",origin)
            dest = paste(dest_city[[i]],dest_province[[i]],sep=",")
            dest = gsub(" ", "+", dest)
            
            ##formatting api request
            directions_Request = paste(host, service,"?origin=",origin,
                               "&destination=",dest,
                               "&key=",api_Key,sep ="")
            
            ## getting nested response list
            raw = GET(directions_Request)
            
            directions_Response = jsonlite::fromJSON(
                content(raw,"text")
            )
            
            ##  Just to check if response is ok
            if(directions_Response$status != "Ok")
                warnings("Request not successful")
            
            
            
            ##geting lat and lng for origin and destination
            
            route_info = directions_Response$routes$legs[[1]]
            
            start_df = route_info[["start_location"]]
            ## a bit truncated
            orig_lat[[i]] = start_df$lat
            orig_lng[[i]] = start_df$lng
            orig_address[[i]] = route_info$start_address
            
            end_df = route_info[["end_location"]]
            dest_lat[[i]] = end_df$lat
            dest_lng[[i]]= end_df$lng
            dest_address[[i]] = route_info$end_address
            
            
            ##getting time, distance info
            time_value[[i]] = route_info$duration$value
            time_text[[i]] = route_info$duration$text
            distance_text[[i]] = route_info$distance$text
            distance_value[[i]] = route_info$distance$value
            
            ##route for map:
            polyline[[i]] = directions_Response$routes$overview_polyline[["points"]]
        }
        
        
        ##making the data frame
        
        df = data.frame(orig_address = unlist(orig_address),
                orig_lat =unlist(orig_lat),
                orig_lng = unlist(orig_lng),
                dest_address = unlist(dest_address),
                dest_lat = unlist(dest_lat),
                dest_lng = unlist(dest_lng),
                time_text = unlist(time_text),
                time_value = unlist(time_value),
                distance_text = unlist(distance_text),
                distance_value = unlist(distance_value),
                polyline = unlist(polyline)
            
        )
        return(df)
        
    }    
```



*   Using the above function to get the data. The data frame
    is saved under "progress2.csv".

```{r testing functions}

    # progress_DF = read.table("progress.txt", header=TRUE,sep=",")
    # test_DF = progress_DF[1:100,]
    # 
    # test= lat_Lng(test_DF$orig_city, 
    #               test_DF$orig_province,
    #               test_DF$dest_city,
    #               test_DF$dest_province,api_Key = api_Key)
    # 
    # if(!file.exists("progress2.csv"))
    #     file.create("progress2.csv")
    # 
    # ##  Some city/destination pairs are unreachable via car only.
    # ##  Thus total orig/destination pairs is less than 100.
    # 
    # test_out = dplyr::select(test, -polyline)
    # write.table(test,"progess2.csv",sep=",",row.names = FALSE)

```


*****


### Calculating Straight Line Distance in km:

*   Internet recommended to use [geosphere](https://cran.r-project.org/web/packages/geosphere/index.html) package.
```{r, message = F, warning=F}
    rm(list = ls())
    require(geosphere)
    progress_DF = read.table("progess2.csv", header=TRUE,sep=",", stringsAsFactors = FALSE)
    
    ## calculating the straight line distance 
    ## for each origin/dest pair
    straight_Dis = list()
    for( i in 1:dim(progress_DF)[1]){
        p1 = as.numeric( c(progress_DF$orig_lng[[i]],progress_DF$orig_lat[[i]]) )
        p2 = as.numeric( c(progress_DF$dest_lng[[i]],progress_DF$dest_lat[[i]]) )
        
        straight_Dis[[i]] = geosphere::distVincentyEllipsoid(p1,p2)/1000
    }
    
    progress_DF = cbind(progress_DF, straight_dis = unlist(straight_Dis))
    

```



*****


### Getting output Data:
*   Saved under "final.csv"

```{r organizing output Data}
    require(dplyr)
    progress_DF = progress_DF %>% mutate(distance_value
        = distance_value/1000) %>%
        mutate( avg_speed = distance_value/time_value*3600) 
    
    out_DF = progress_DF %>%
        select(-c(polyline,time_value,distance_text))
    
    write.csv(out_DF,"final.csv",
        col.names = TRUE, row.names = FALSE)
        

```



*****


### Decoding polylines:
*   Decoding Polyline Function copied from
    [here](http://s4rdd.blogspot.com/2012/12/google-maps-api-decoding-polylines-for.html)

*   [More info on polylines](https://developers.google.com/maps/documentation/utilities/polylinealgorithm)

```{r}
# Custom decode function
# Taken from http://s4rdd.blogspot.com
# /2012/12/google-maps-api-decoding-polylines-for.html

decodeLine <- function(encoded){
  require(bitops)

  vlen <- nchar(encoded)
  vindex <- 0
  varray <- NULL
  vlat <- 0
  vlng <- 0

  while(vindex < vlen){
    vb <- NULL
    vshift <- 0
    vresult <- 0
    repeat{
      if(vindex + 1 <= vlen){
        vindex <- vindex + 1
        vb <- as.integer(charToRaw(substr(encoded, vindex, vindex))) - 63  
      }

      vresult <- bitOr(vresult, bitShiftL(bitAnd(vb, 31), vshift))
      vshift <- vshift + 5
      if(vb < 32) break
    }

    dlat <- ifelse(
      bitAnd(vresult, 1)
      , -(bitShiftR(vresult, 1)+1)
      , bitShiftR(vresult, 1)
    )
    vlat <- vlat + dlat

    vshift <- 0
    vresult <- 0
    repeat{
      if(vindex + 1 <= vlen) {
        vindex <- vindex+1
        vb <- as.integer(charToRaw(substr(encoded, vindex, vindex))) - 63        
      }

      vresult <- bitOr(vresult, bitShiftL(bitAnd(vb, 31), vshift))
      vshift <- vshift + 5
      if(vb < 32) break
    }

    dlng <- ifelse(
      bitAnd(vresult, 1)
      , -(bitShiftR(vresult, 1)+1)
      , bitShiftR(vresult, 1)
    )
    vlng <- vlng + dlng

    varray <- rbind(varray, c(vlat * 1e-5, vlng * 1e-5))
  }
  coords <- data.frame(varray)
  names(coords) <- c("lat", "lon")
  coords
}


```



*****


### Setting up Basic Map

```{r basic map, message = F, warning=F}
    library(leaflet)
    library(htmltools)
    
    ##calculating max/min lat and lng for the bounding box.
    lat = append(progress_DF$orig_lat, progress_DF$dest_lat)
    min_Lat = min(lat)
    max_Lat = max(lat)
   
    lng = append(progress_DF$dest_lng, progress_DF$orig_lng)
    min_Lng = min(lng)
    max_Lng = max(lng)


    #   Making Icons
    orig = makeAwesomeIcon(markerColor="red", iconColor="yellow")
    
    # adding popup labels
    popup = function(address){
        list = lapply( address, function(x){
            split = unlist( strsplit(x, ",") )
            city = paste("city:", split[1])
            province = paste("province: ", split[2])
            country = paste("country: ", split[3])
            out = paste(city, province, country, sep = "<br/>")
        })
        
        return( as.character(list) )
    }
  
    
   
    ##  loading basic map. 
    map = leaflet() %>%
        addProviderTiles(provider="OpenStreetMap.Mapnik") %>%
        fitBounds(max_Lng,max_Lat,min_Lng,max_Lat) %>%
    ##  adding markers
        addAwesomeMarkers(lng=progress_DF$orig_lng,
                   lat=progress_DF$orig_lat,
                   icon=orig,
                   popup = popup(progress_DF$orig_address)) %>%
        addMarkers(lng=progress_DF$dest_lng,
                   lat=progress_DF$dest_lat) %>%
     ##  Adding legend: 
        addLegend(position = "topright", 
            labels = c("origin","destination"),
            color = c("Red","Blue"), layerId = "legends")
    ## adding Polyline
      
    for(i in 1: length(progress_DF$polyline)){
        df = decodeLine(as.character(progress_DF$polyline[[i]]))
        
        map = map %>%
            addPolylines(lng=df$lon,lat=df$lat, color = "yellow")
    }
    
   
    
    
    map
    
```
