---
title: "Google Visualizing"
author: "paul"
date: "March 8, 2017"
output: html_document:
        toc: true
        toc.depth: 4
---

```{r googleVisualizingSetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*****

### Overview:

*   There are many packages in R to help with visualization of geographical 
    location data. The following is a few of the useful packages:
    +   _[ggplot2](http://ggplot2.org/):_ provides an interface to simplify plotting of 
        data and helps to visualize geographical data. 
    +   _[tmap](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-nutshell.html):_ Allows flexible and interactive functions to 
        produce thematic maps. 
    +   _[ggmap](https://blog.dominodatalab.com/geographic-visualization-with-rs-ggmaps/):_ Provides an easy way to overlay static maps 
        from google Maps and other sites to visualize data. 
    +   _[leaflet](https://rstudio.github.io/leaflet/):_ R version of the popular JavaScript library.
    This is used to create interactive maps.
    +   _[SP Package](https://cran.r-project.org/web/packages/sp/vignettes/intro_sp.pdf):_ 
        This package is one of the fundamental packages in R that deals with spatial data. 
        It was observed that many of the packages references back to the SP package.

*****

### Leaflet:
*   The Visualization section of the report will mainly focus on using 
    the leaflet package. It appears that leaflet package offers the most
    flexibility and customizability for visualization. 

*   The down side is that this can only be loaded via connection 
    to the internet. 
    
*   Since this package works with html and javascript in it's underlying structure.
    The interactive maps can not be displayed in PDF format. 
    
*****

#### Example 1: Online Tutorial
*   [Introduction videos to Leaflet](https://www.youtube.com/watch?v=ktWjjAlAKtE)

```{r VisualizationPackages, message = F}
    rm(list = ls());

    library(leaflet)
    
```

```{r Visualization Example 1.1 }

    ## "leaflet()" essentially initializes a window to display the map
    ## addTiles() allows placing layers
    mymap = leaflet() %>%
        addTiles()
    
    
    ## Use tile providers to include custom map types
    mymap = addProviderTiles(mymap,provider="OpenStreetMap.Mapnik")
    mymap
```
*   __[Preview of Other Map Providers](https://leaflet-extras.github.io/leaflet-providers/preview/)__

```{r Visualization Example 1.2}
    ## Seting view and adding markers
    
    taj_Mahal_Lat = 27.1750
    taj_Mahal_Lng = 78.0419
    ## SetView allows user to set view based on a center coordinate
    mymap = setView(mymap, lng=taj_Mahal_Lng, lat=taj_Mahal_Lat, zoom=15)
    
    ## Adding a marker at Taj Mahal
    mymap = addMarkers(mymap,lng=taj_Mahal_Lng,lat=taj_Mahal_Lat,
                       popup="Taj Mahal")
    mymap
    rm(list = ls())
```



*****

### Example 2: Showing Some Google Map Outputs:

*   __Goal: Tracing the directions from Work to an apartment, showing places of interest
      around these locations with appropriate labels.__

*****

#### Functions
*   [Polyline](https://developers.google.com/maps/documentation/utilities/polylinealgorithm)
    +   A polyline is a list of lat/lng pairs compressed to a string via converting 
        the bytes into a string.


*   Rather useful for unpacking JSON object
    [tidyjson](https://cran.r-project.org/web/packages/tidyjson/vignettes/introduction-to-tidyjson.html)

*   Tidyjson is useful if the nest list has multiple elements with the same names. 
    However,for single and small bits of information, direct subsetting using "[[  ]]"
    will be the easiest. 


*   A function used to extract lat/lng information
    for the origin and destination of the directions response,
    using tidyjson.
    
```{r}

    extracting_Origin_End = function(response_Text)
{   ##This is to illustrate how to get the lat/lng using the tidyjson format.
        ## Consider the tbl_json object as a variable data frame
        ## where dimensions are constantly changed by functions such as
        ## spread, stack
        
        
    require(tidyjson)
    require(dplyr)


    # Extracting direction markers for start: 
    directions_Markers_Start =response_Text %>% as.tbl_json %>% 
        enter_object("routes") %>% gather_array %>%
        enter_object("legs") %>% gather_array %>%
        spread_values(
            address = jstring("start_address")
        ) %>% enter_object("start_location")%>%
        spread_values(
            lat = jnumber("lat"),
            lng = jnumber("lng")
        ) %>% select(address, lat,lng)%>% 
        tbl_df
    
    directions_Markers_End = response_Text %>% as.tbl_json %>% 
        enter_object("routes") %>% gather_array %>%
        enter_object("legs") %>% gather_array %>%
        spread_values(
            address = jstring("end_address")
        ) %>% enter_object("end_location")%>%
        spread_values(
            lat = jnumber("lat"),
            lng = jnumber("lng")
        ) %>% select(address, lat,lng)%>% 
        tbl_df 
        
        
    directions_Markers = rbind(directions_Markers_Start,directions_Markers_End)
    return(directions_Markers)
}
    
```


*   Decoding Polyline Function borrowed:

```{r}
# Custom decode function
# Taken from http://s4rdd.blogspot.com
# /2012/12/google-maps-api-decoding-polylines-for.html

decodeLine <- function(encoded){
  require(bitops)

  vlen <- nchar(encoded)
  vindex <- 0
  varray <- NULL
  vlat <- 0
  vlng <- 0

  while(vindex < vlen){
    vb <- NULL
    vshift <- 0
    vresult <- 0
    repeat{
      if(vindex + 1 <= vlen){
        vindex <- vindex + 1
        vb <- as.integer(charToRaw(substr(encoded, vindex, vindex))) - 63  
      }

      vresult <- bitOr(vresult, bitShiftL(bitAnd(vb, 31), vshift))
      vshift <- vshift + 5
      if(vb < 32) break
    }

    dlat <- ifelse(
      bitAnd(vresult, 1)
      , -(bitShiftR(vresult, 1)+1)
      , bitShiftR(vresult, 1)
    )
    vlat <- vlat + dlat

    vshift <- 0
    vresult <- 0
    repeat{
      if(vindex + 1 <= vlen) {
        vindex <- vindex+1
        vb <- as.integer(charToRaw(substr(encoded, vindex, vindex))) - 63        
      }

      vresult <- bitOr(vresult, bitShiftL(bitAnd(vb, 31), vshift))
      vshift <- vshift + 5
      if(vb < 32) break
    }

    dlng <- ifelse(
      bitAnd(vresult, 1)
      , -(bitShiftR(vresult, 1)+1)
      , bitShiftR(vresult, 1)
    )
    vlng <- vlng + dlng

    varray <- rbind(varray, c(vlat * 1e-5, vlng * 1e-5))
  }
  coords <- data.frame(varray)
  names(coords) <- c("lat", "lon")
  coords
}


```
   
   
*****

#### Main Example:  

```{r Visualization Example 2 Libraries, warning=F,message=F}
 # Libraries:
    library(leaflet)
    library(httr)
    library(jsonlite)
    library(tidyjson)
    library(dplyr)
    library(bitops)

```


*   Making directions request and getting response data:

```{r visualization Example 2 Directions Data, warning=F}

   

    # Setting up Directions requests:
    apartment_Address = "147+Millgreen+Crescent,+Gloucester"
    work_Address = "M-50,+1200+Montreal+Rd,+Ottawa,+ON+K1A+0R6"
    api_Key = "AIzaSyBGTs-gZCbyP8n0Hvw_VZ76Z6YrST1DNa8"
    host = "https://maps.googleapis.com/maps/api"
    
    directions_Request = paste(host, "/directions/json","?origin=",work_Address,
                               "&destination=",apartment_Address,"&mode=walking",
                               "&key=",api_Key,sep = "")
    
    # Extracting directions data
    raw = GET(directions_Request)
    
    directions_Response = jsonlite::fromJSON(
        content(raw,"text")
    )
    
    
    # Parsing route information
    directions_Markers = extracting_Origin_End(content(raw,"text"))
        ## The route will be a rough polyline 
        ## extracted from the "overViewPolyline"
    route = decodeLine(
        directions_Response$routes$overview_polyline[["points"]]
    )
    
    ## The main diffculty is that you have to know JSON format
    ## and thus it will take trials to parse the information.
    
    route_Information = data.frame(
        distance = directions_Response$routes$legs
        [[1]][["distance"]][["text"]],
        duration= directions_Response$routes$legs
        [[1]][["duration"]][["text"]]
    )
     
     rm(list = c("directions_Request",
                 "raw", "directions_Response"))
    
```


*   Plotting the directions on a map:

```{r Visualization Example 2 Plotting Directions,warning=F}

    #loading basic map. 
    example2_Map = leaflet() %>%
        addProviderTiles(provider="OpenStreetMap.Mapnik") %>%
        fitBounds(lng1=directions_Markers$lng[[1]],
                  lat1=directions_Markers$lat[[1]],
                  lng2=directions_Markers$lng[[2]],
                  lat2=directions_Markers$lat[[2]]
                  )
    
    # Adding Markers: 
    example2_Map = example2_Map %>%
        addMarkers(lng=directions_Markers$lng, 
                   lat=directions_Markers$lat, 
                   label=c("work","home")
        )
    
    #Adding route information:
    popup_Content = paste(sep = "","distance: ",
                          route_Information$distance,"<br>",
                          "duration: ", route_Information$duration,
                          "<br>")
    
    example2_Map = example2_Map %>%
        addPolylines( lng=route$lon, 
            lat=route$lat,
            popup=popup_Content
        )
    
    example2_Map
    
   
```



*   Adding Places of Interest:

```{r Visualization Example 2 Places_Of_Interest}

    # Setting up for URL requests
    work_Lat = directions_Markers$lat[[1]]
    work_Lng = directions_Markers$lng[[1]]
    home_Lat = directions_Markers$lat[[2]]
    home_Lng = directions_Markers$lng[[2]]
    keyword = "point+of+interest"
    ## circular radius in meters
    radius = 2000
    
    
    #function to extract rawData:
    place_Raw_data = function(lat, lng, keyword, radius, api_Key)
    {
        request = paste(host,"/place/nearbysearch/json",
        "?location=",lat,",",lng,"&radius=",
        radius,"&keyword=",keyword, "&key=",
        api_Key,sep="")
        
        raw = jsonlite::fromJSON(content(
                GET(request),"text"
            ))
        
    }
    
    # Getting Raw Data
    poi_Work_Raw= place_Raw_data(work_Lat, work_Lng,
                            keyword, radius, api_Key)
    poi_Home_Raw = place_Raw_data(home_Lat, home_Lng,
                            keyword, radius, api_Key)
    
    # Function to format Raw Data into organized DF:
    
    organizePlaceDF= function(raw_Data)
    {
        lat_Lng = raw_Data$results$geometry$location
        names = as.vector(unlist(raw_Data$results$name))
        icon = as.vector(unlist(raw_Data$results$icon))
        rating = as.vector(unlist(raw_Data$results$rating))
        organized_Data = cbind(lat_Lng,rating,icon,names)
        organized_Data
    }
    
    
    # getting DF
    home_Df = organizePlaceDF(poi_Home_Raw)
    work_Df = organizePlaceDF(poi_Work_Raw)
    place_Df = rbind(home_Df,work_Df)
    
    #setting up Icons:
    place_Icons <- icons(
      iconUrl = place_Df$icon,
      iconWidth = 20, iconHeight = 20,
      iconAnchorX = 0, iconAnchorY =0
    )
    
    #setting up label contents:
    content =list()
    for(i in 1:length(place_Df$icon))
    {
        content_Name= paste("name:",
            place_Df$names[[i]],"<br>")
        
        content_Rating= paste("rating:",
            place_Df$rating[[i]],"<br>")
        
        content[[i]] = 
            paste(content_Name, content_Rating,
                  sep="<br>")
        
    }
    
    
    # adding the locations onto the map using google Icons
    example2_Map = addMarkers(example2_Map, lat=place_Df$lat, 
                    lng=place_Df$lng,icon=place_Icons,
                    popup=as.vector(content))
    
    
    
     example2_Map
    
```





*****



### Example 3: Visualizing Density
__GOAL: The goal is to map the number of crimes in Alabama  
as a percentage of population and display it
on the map.__


*  Downloading and Organizing Data:
```{r Visualitzation Example 3, warning=F, message=F}
    rm(list = ls())
    require(httr)
    require(dplyr)
    require(leaflet)
    require(jsonlite)

    url = "https://ucr.fbi.gov/crime-in-the-u.s/2013/crime-in-the-u.s.-2013/tables/table-8/table-8-state-cuts/table_8_offenses_known_to_law_enforcement_alabama_by_city_2013.xls/output.xls"
    api_Key = "&key=AIzaSyBGTs-gZCbyP8n0Hvw_VZ76Z6YrST1DNa8"
    host = "https://maps.googleapis.com/maps/api/geocode/json"
    query = "?address="
    
    
##  Could not figure out why the file can not be opened.
##  It is manually downloaded and reformated to a 
##  excel format.

    crime_Raw_Data = read.csv(file = "./USA_Crime_Data.csv",
                          header=T,stringsAsFactors = F)

    ## Reformatting Df to contain (city,population, total crime number)
    
    ## Sorting through NA = 0
    for(i in names(select(crime_Raw_Data,
            -(City))))
    {
        list = crime_Raw_Data[,i]
        list = as.numeric(list)
        list[is.na(list)] = 0
        crime_Raw_Data[,i] = list

    }
    
    numeric_Data = select(
            crime_Raw_Data,
            Violent.crime:Arson3
        )
    
    ## new data frame
    data = cbind(
        city=crime_Raw_Data$City,
        fraction=( rowSums(numeric_Data)/
            as.numeric(crime_Raw_Data$Population))
    )
    # rm(list = c("numeric_Data","crime_Raw_Data"))
    ```
    
    
*   Organizing Lat/Long : 


```{r Visualization Example3: Lat and Lng}
    ## function to extract lat and long for a list of cities
    ## in USA
    ##only the first 50 were plotted, it takes quite a while
    ## to retrieve the data otherwise. 

     length = 50
     list = data[,"city"][1:length]
     
     location_Data = lapply(list,function(x){
         
                 x = gsub(pattern=" ", replacement= "+",
                      x=x)
                 request = paste(host, query, x,
                            ",USA",api_Key,sep="")
                 
                 raw = jsonlite::fromJSON(
                     content(GET(request),"text")
                 )
                raw$results$geometry$location  
                
                ##  The results for 19th element
                ## "Bear creek" returned multiple
                ## elements.Dont know why
        })
     
   lat = list()
   lng = list()
    
    ## merging the lat and lng into the data DF
    for(i in 1:length)
    {
        ##just going to ignore the 19th element for now
           lat[[i]]= location_Data[[i]]$lat
            lng[[i]]= location_Data[[i]]$lng 
            
        if(i==19)
        {
            ##Selecting correct lat data
            string = toString(unlist(
                location_Data[[i]]$lat
            ))
            
            index = regexpr(
                pattern=",",
                text=string
            )
            
            lat[[i]]=as.numeric(
                substr(
                    string,1,
                    as.numeric(index[1])-1
                )
            )
            
            
            ##Selecting correct lng data
            string = toString(unlist(
                location_Data[[i]]$lng
            ))
            
            index = regexpr(
                pattern=",",
                text=string
            )
            
            lng[[i]]=as.numeric(
                substr(
                    string,1,
                    as.numeric(index[1])-1
                )
            )
            
            
        }
    }
```
   
  
*   Organizing information needed for the map:

```{r Visualization Example3: Map data}  
    org_Data = cbind(data[1:length,1:2],lat,
            lng)
    org_Data = as.data.frame(org_Data)
   
     ## Reorganizing data frame to change to numeric
    org_Data$fraction = as.numeric(org_Data$fraction)
    org_Data$lat = as.numeric(org_Data$lat)
    org_Data$lng = as.numeric(org_Data$lng)
    
    
    ##popup labels:  
    popup = list()
    for(i in 1:length(org_Data$fraction)){
        precentage = paste("Precentage:",
            signif(org_Data$fraction[[i]],4)*100)    
        
        city = paste("City:",org_Data$city[[i]])
        
        return = paste(city, precentage,sep="<br>")
        popup[i] = return
        
    }
    
    
    
   
    
```


*   Plotting the data onto a map:


```{r Visualization Example3: Map}
    ##setting up the Map
    center_Lat = median(org_Data$lat)
    center_Lng = median(org_Data$lng)
    
    example3_Map = leaflet(data=org_Data) %>% 
        addProviderTiles("OpenStreetMap.DE")%>%
        setView(lat=center_Lat,lng=center_Lng,
                zoom = 8) 
        
    
  
    
    
    ## Adding circle markers proportional to 
    ## the fraction.
    radius_Factor = 300
    
     example3_Map = example3_Map %>%
         addCircleMarkers(lng=org_Data$lng,
             lat=org_Data$lat,
             popup=popup,
             radius = org_Data$fraction*radius_Factor)
    
    
        example3_Map  
```